---
title: "fhiRclient"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fhiRclient}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction
The R package is developed to implement the FHIR client with essential features of authentication, CRUD operations, bundle search, and data structures for the FHIR resources. The package is built based on the python library `fhircient` and S4 class. Tidy-style functions were developed for the clinial data manipulation more friendly.

# Import the package
To get started, first we need to load the package.
```{r setup}
library(fhiRclient)
library(dplyr)
```

The dependent python `fhirclient` can be installed from the build-in method. The latest version can be installed from its github repository directly for the version of R4 support.
```{r, eval=FALSE}
install_fhirclient(version = "latest")
```

# Building connection
```{r}
ct <- Client(app_id = "my_app",
             api_base = "https://r4.smarthealthit.org")
ct
```

# CRUD functions
## Read a Patient record
To read a patient record and convert it to json list.
```{r}
pt <- Patient(list(id = "326b4675-0bc8-4dbd-b406-a5564c282401"))
ct %>% Read(pt) %>% as_json %>% as_json_tbl
```

To read a patient and get the birth date.
```{r}
ct %>% Read(pt) %>% BirthDate
```

To read a patient and get the full name.
```{r}
ct %>% Read(pt) %>% getName
```

## Create a new Patient record
```{r}
pt1 <- Read(ct, pt)
pt1$id <- NULL
newpt <- ct %>% Create(pt1)
newID <- newpt$id
newID
```

## Update a patient record
An example to add middle name and update the record.
```{r}
pt1 <- ct %>% Read(Patient(list(id = newID)))
pt1 %>% getName
pt1$name[[1]]$given <- list("Bradly", "Middle")
update <- ct %>% Update(pt1)
ct %>% Read(pt1) %>% getName
```

## Delete a patient record
```{r}
ct %>% Delete(pt1)
```

# Bundle search
The `Search` function can be used to search for resource bundles with parameters. The `page` option can be specified to look up results from first page or all pages.
```{r}
res <- Search(ct, "observation", "Observation",
              list(code = "http://loinc.org|2339-0"), page = "all")
res
```

The `filterBundle` function can be used to extract specific elements. For example, here we extract all the unique reference patients.
```{r}
res %>% filterBundle('subject') %>% unique
```

We can group the results with indexes and convert it to wide table with patients in different columns.
```{r}
gres <- res %>% filterBundle("entry.resource") %>% group_index("entry.resource.id")
gres
gres[1:44,] %>% tidyr::pivot_wider(names_from = entry.resource.id)
```

# Data Classes
Here we have two ways to create a patient model, with `Patient` functon or a general `Model` constructor. 
```{r}
pt <- Patient(list(id = "pid"))
pt <- Model("patient", "Patient", list(id = "pid"))
```

The function `getModelClass` can be used to query all available classes for a model. The general `Model` function require the resource model name and a data class to build.
```{r}
getModelClass("humanname")
hn <- Model("humanname", "HumanName", list(family = "A", given = list("B", "C")))
hn
as_json(hn)
```

# Shiny Demo
```{r, eval=FALSE}
ShinyDemo()
```

# SessionInfo
```{r}
sessionInfo()
```
